\myphantomsection{Acknowledgments}{

% \chapter* does not set a head mark!
\addchap*{Acknowledgments}

This dissertation represents more than just a piece of writing, it is a milestone in years of work at the University of Victoria. This work was possible thanks to the collaboration and support of many people, whom I want to acknowledge and thank.

I would first like to express my utmost gratitude to my supervisors, professors Hausi A. M\"{u}ller and Gabriel Tamura. Their continuous guidance and support throughout this journey, and their motivation and inspiration to endure made this work possible. I will be forever grateful for all their teachings during this time. Thank you, Hausi and Gabriel, for your patience during our meetings and brainstorming sessions. Thank you for your ingenious suggestions to improve my manuscripts. I was lucky to have you both as my supervisors. Your distinct perspectives were always insightful. I will not forget the words of encouragement, the \emph{have fun} and the Fridays to think big thoughts.

I would also like to extend my deepest gratitude to professor Norha M. Villegas, who provided invaluable advice and contributions into my work. Her extensive knowledge, constructive criticism and guidance had a great impact on my publications and on my formative process. Norha, I greatly appreciate the practical and insightful suggestions you provided. I will not forget the happy faces and encouragement phrases you always added to your annotated reviews.

I would like to express my deepest appreciation to my committee, professors Neil Ernst and Issa Traor\'{e}, for their insightful comments and hard questions. I also thank Dr. Grace Lewis for acting as the external examiner of my dissertation, and professor Richard Marcy, for serving as the chair of my final oral examination.

To the members of the Rigi-Pita team, Ulrike, Lorena, Felipe, Priya, Ivan, Kirti, Giovanni, Sunil, Karan and Alvi, thank you for the meaningful, and meaningless, conversations we had in the lab. I appreciate and cherish my memories of us spending time and bonding at CASCON, CASTLE and CSER, and will not forget the sleepless nights we worked together before deadlines. I believe the Rigi lab will have the perfect coffee machine someday. Felipe, it has been a pleasure working with you.

I would like to acknowledge the institutions that sponsored my research. This dissertation, and the publications derived from it, were possible thanks to the funding provided by the University of Victoria, the National Sciences and Engineering Research Council (NSERC) of Canada, IBM Corporation, and Universidad Icesi (Colombia). I am specially thankful to the Center for Advance Studies (CAS) at IBM for the opportunity of being a CAS student in multiple projects. Particularly helpful to me during this time were Joe Wigglesworth, Ian Watts, Hugh Hockett, Joanna W. Ng, and Vio Onut for their valuable feedback, guidance and support.

I would like to express my gratitude to Joe Wigglesworth, Ian Watts, and Hugh Hockett. Thank you for helping shape our contributions and for your guidance and patience during the patenting process of our invention. Each of you contributed in many ways that I appreciate.

I would also like to extend my sincere thanks to the faculty and administrative staff of the University of Victoria. Thank you Kath Milinazzo, Nancy Chan, Wendy Beggs, and Erin Robinson, you have contributed in many ways to the completion of my research. Nancy, your patience cannot be overestimated. I would also like to extend my gratitude to the faculty of the computer science department, especially professors Ulrike Stege, Sudhakar Ganti and Daniela Damian. Thank you for giving me advice, reading manuscripts and enriching my academic experience. Ulrike, I appreciate your constant support at our group meetings. You were one of the first people with whom I worked when I started my PhD, I remember you were very supportive, kind and patient.

I also wish to thank all the researchers whom I have met many times during academic engagements. Among many others, I would like to thank professors Kelly Lyons, Kostas Kontogiannis, Marin Litoiu, Marios-Eleftherios Fokaefs, and Nelly Bencomo for the stimulating discussions we held offline and online, and their feedback at conferences and other academic events. I am grateful for the training, networking and participation events of the Dependable Internet of Things Applications (DITA) program. Many thanks to Shabnam Nikfar for coordinating the DITA training sessions and student conferences. Lastly, I would like to acknowledge students from across Canada with whom I shared joyful moments during long conference days: Andr\'{e}s Paz, Cristiano Politowski, Fabio Petrillo, Hao Jiang, Manel Grichi, Marios-Stavros Grigoriou, Mohab Aly, Mouna Abidi, Zeinab kermansaravi, and many others. I cherish these memories and thank you all for being a part of my student life in Canada.

The completion of this thesis would not have been possible without the unparalleled support and nurturing of Juanita. Thank you, for always believing in me. Thank you for lifting me up when the going was rough. Your company through this journey kept me strong. I am deeply indebted to you.

I am lucky to have a group of wonderful friends. I want to thank you, Lorena, Carlos, David, Geo, Maryi, and Felipe, for the best moments I had in Victoria. I hold dear every memory of us spending time, and wish we can repeat them sometime. I cannot think of my friends without mentioning Jessica and Fabi\'{a}n. From a distance, you accompanied me through every step of this journey, and helped me stay focused. I treasure your friendship, and very much appreciate our long-distance calls, our plans, and everything. To my friends in Cali, Diego, Cristiam, Jeffer, Alexis and Danny, I thank you for the good times, and for always keeping an eye on me.

I end this section by talking about Colombia, where the people I love and miss the most reside: my family. Their support has been unconditional all these years. They have been by my side from a distance, each in their own unique ways. I thank them in Spanish now.

\emph{¡Gracias a todos! A mi abuelita Marta, quien se nos fue mientras hac\'{i}a el doctorado, le agradezco por ense\~{n}arme el valor de la humildad. A mi mam\'{a} y a mi pap\'{a}, Amelia y Nelson, les agradezco el apoyo incondicional y la confianza que desde siempre han depositado en mi. A mi hermano Andr\'{e}s le agradezco por darme motivos para terminar. A mi padrino Fredy le agradezco las llamadas a larga distancia, siempre cargadas de noticias y hechos cient\'{i}ficos; me las he gozado de principio a fin. A mi t\'{i}a Teresa le agradezco por acompañarme desde Francia; su compa\~{n}\'{i}a ha sido una gu\'{i}a para mis experiencias en Canad\'{a}. Siempre estar\'{e} agradecido con mi t\'{i}o Helbert por haberme introducido al mundo de los sistemas; a pesar de nuestra distancia de un tiempo a esta parte, sus ense\~{n}anzas no caducan. A mis t\'{i}as y a mis t\'{i}os les agradezco los buenos deseos y las bendiciones cada vez que hablamos.}

}% Phanton section
